---
title: "Design of Data Analysis Routines"
subtitle: "A Design Process Description"
author: "Peter von Rohr"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

# Disclaimer
This document is the result of a refactoring process of the original version of the QA and AM document [@vonRohr2018]. This document is focusing on the design of data analysis routines where as in the new version of the QA and AM document [@vonRohr2018] contains only the parts on quality assurance and automation. This allows for a better structuring and a clearer separation of the different concepts. 


# Introduction
One main area of activity for breeding organisation are genetic evaluation data analysis routines. Such routine analyses are done according to a given schedule. Depending on the livestock species this might happen on a daily basis or only a few times a year.


# Terminology



```{r bib, include=FALSE}
vec_bref <- c(
  bibentry(
    bibtype = "Unpublished",
    title   = "Concepts for QA and AM",
    author  = "Peter von Rohr",
    year    = 2018,
    url     = "<insert URL to Confluence page>",
    note    = "An internal document on QA and AM of data analysis processes",
    key     = "vonRohr2018"))

# write bib file
rmddochelper::write_bib(pvec_bref = vec_bref)
```
